### 2013-07-11 monte_carlo.tex
Copied whole sections from the phd.tex.

\documentclass[article.tex]{subfiles} 
\begin{document}

\section{Monte Carlo Simulations}
\label{sec:monte-carlo-simul}

\todo[inline]{Write the Monte-Carlo section}

\end{document}
% -----------------------------------------------------------