### 2013-07-27 iccp2013.tex
added the iccp2013 article

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{iccp}
\usepackage{times}
\usepackage{graphicx}
\graphicspath{{./images/}}
\usepackage{epstopdf}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}
\usepackage{bm}
\usepackage{enumerate}

%
% Some new commands I use in this text
%
\newcommand{\OpSphere}{\bm{\mathcal{S}}}
\newcommand{\OpRot}{\bm{\mathcal{R}}}
\newcommand{\OpDistance}{\bm{\mathcal{D}}}
\newcommand{\OpCumsum}{\bm{\mathcal{C}}}
\newcommand{\OpInt}{\bm{\mathcal{I}}}
\newcommand{\OpCamera}{\bm{\mathcal{P}}}
\newcommand{\OpDiag}[1]{\mathrm{diag}\left\{#1\right\}}
\newcommand{\Grad}[1]{\bm{\triangledown_{#1}}}
\newcommand{\argmin}{\mathrm{arg}\min}
\newcommand{\curly}[1]{\left\{#1\right\}}
\newcommand{\roundy}[1]{\left(#1\right)}
\newcommand{\recty}[1]{\left[#1\right]}
\newcommand{\PartDeriv}[2]{\frac{\partial{#1}}{\partial{#2}}}
\newcommand{\vect}[1]{\bm{#1}}
\newcommand{\mat}[1]{\bm{#1}}
\newcommand{\transpose}[1]{{#1}^\intercal}
\newcommand{\derivsym}[1]{\,d{#1}}

\newcommand{\yoavcomment}[1]{}
\renewcommand{\yoavcomment}[1]{#1} % Comment to remove images

%\iccpfinalcopy % *** Uncomment this line for the final submission

\def\iccpPaperID{19} % *** Enter the ICCP Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ificcpfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{Sky Tomography for 3D Aerosol Distribution Recovery}

\author{Amit~Aides\\
  Electrical Engineering Dept.\\
  Technion - Israel Inst. Tech.\\
  Haifa 32000, Israel\\
  {\tt\small amitibo@tx.technion.ac.il} \and
  Yoav~Y.~Schechner\\
  Electrical Engineering Dept.\\
  Technion - Israel Inst. Tech.\\
  Haifa 32000, Israel\\
  {\tt\small yoav@ee.technion.ac.il} \and
  Michael~J.~Garay\\
  Jet Propulsion Laboratory\\
  California Inst. Tech.\\
  CA 91011, United States\\
  {\tt\small Michael.J.Garay@jpl.nasa.gov}
  % \and
  % Vadim~Holodovsky\\
  % Electrical Engineering Dept.\\
  % Technion - Israel Inst. Tech.\\
  % Haifa 32000, Israel\\
  % {\tt\small vholod@tx.technion.ac.il}
}

\maketitle
\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
  Aerosols effect climate, health and aviation.  Currently, their
  retrieval assumes a plane-parallel atmosphere and solely vertical
  radiative transfer. We propose a principle to recover the atmosphere
  as it really is: a three dimensional (3D), volumetric, transparent
  huge object. Tomography is the key to recovery. The process involves
  acquiring the sky lightfield by multi-view wide-angle photography on
  a very large scale. The tomography model is distinct, as the light
  source is unidirectional and uncontrolled, while off-axis scattering
  dominates the images.  We formulate an image formation model based
  on 3D radiative transfer. Model inversion is done using optimization
  methods, exploiting a closed-form gradient which we derive for the
  model-fit cost function. The problem is complex but tractable and
  parallelizable, eventually estimating the 3D aerosol distribution.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}
\label{sec:intro}

Lightfield and integral
imaging~\cite{bishop,horstmeyer,kim,levoy,Perwass} samples the optical
radiance distribution in location and direction. It is mainly used in
small-scale setups. This paper deals with a huge scale, up to Earth's
size.  The atmosphere allows light to pass through in multiple
locations and directions. Light is affected by this medium. Therefore,
in this paper we lay out a principle to recover this 3D medium using
measured and modeled lightfields (Fig.~\ref{fig:front}).
\begin{figure}[t!]
  \centering
  \yoavcomment{\includegraphics[width=\columnwidth]{images/front_img.pdf}}
  \caption{\small The sky is photographed from multiple viewpoints
    using wide angle (hemispherical) cameras, in a very wide area. The
    three dimensional distribution of aerosols in the atmosphere
    affects the images. This distribution is then recovered by a
    dedicated tomography formulation.}
  \label{fig:front}
\end{figure}

Projecting this large object to various directions is similar to
tomography in other scientific domains. However, the formulation here
is distinct. Most tomography setups have a controlled and/or
multidirectional radiation source~\cite{gorbunov,messer}. In contrast,
our source is the uncontrolled, unidirectional Sun. Moreover, typical
tomography relies on a linear model~\cite{gregson}: the pixel value is
a linear combination of components along a line of sight (LOS), or a
multiplicative combination (linearized by a logarithm). Linear
tomography can detect volcanic gas~\cite{wright}, which absorbs
UV. Other gases in the atmosphere absorb IR. However, in visible
light, atmospheric attenuation is typically dominated by {\em
  scattering} by particles (aerosols), rather than absorption by
gasses. Thus, in our problem, we cannot rely on direct illumination,
but on sunlight scattered into the LOS~\cite{Seidel}.  The model turns
out to be {\em nonlinear, yet tractable}.

We do not deal with clouds here. Clouds scatter sunlight very strongly
and are easily observed. Hence they are directly recovered using
common stereoscopic tools, similarly to opaque surfaces~\cite{seiz}.
On the other hand, aerosols have much more subtle visual effects and
are thus more difficult to recover. We focus on them.  In remote
sensing~\cite{diner}, imaging through air is associated with {\em
  atmospheric correction}, based on {\em aerosol retrieval}. In this
discipline, the atmosphere is often assumed to be plane-parallel, for
simplicity, using only 1D vertical radiative transfer. Consequently
state-of-the-art aerosol retrieval is done in distinct large lateral
blocks~\cite{matronchik} with limited height
resolution~\cite{kalashnikova}. We, however, seek high resolution 3D
recovery. Computational photography offers new ways to peer into
media. In addition to active
descattering~\cite{fuchs,kim,levoy,Marschner} and recovery of
refractive transparent objects~\cite{ihrke, Saito}, overcoming the
atmosphere~\cite{Joshi2,Milanfar} to view ground objects was studied
in horizontal, fixed
views~\cite{fattal,he,kopf,kratz,narasimhan2,oakley,Hschechner2,tan}. In
this paper, however, the medium itself, at all relevant altitudes is
the object of interest.

3D recovery of this medium has direct implications to various
scientific communities that either rely on remotely-sensed imagery,
study the atmosphere, or overcome the medium to see beyond. These
include meteorology, atmospheric sciences, volcanology, and
climatology.  Aerosol retrieval is important for understanding climate
evolution~\cite{dayan,kalashnikova} and monitoring air
quality. Mapping aerosol density is significant to aviation safety,
which needs real time assessment of conditions and visibility around
flight paths. To keep safe, today's coarse sensing limits flight
capacity. A prominent example is the 2010 eruption of Iceland's
Eyjafjallaj$\ddot {\rm o}$kull volcano, whose ash (aerosols) covered
much of Europe's airspace, grounding $\approx 100,000$
flights. Knowing the 3D distribution of aerosol density may indicate
safe 3D corridors to safely pass through, enhancing airspace
use. Viewing beyond the medium, atmospheric correction is needed, for
example, in satellite tracking of chlorophyll
distributions~\cite{johnsen}. Additional bodies in the solar system
have atmospheres. These objects can be viewed from multiple directions
by probes.  The approach described here may be helpful there.

We model passive optical tomographic imaging of 3D atmospheric
scatterer distributions in cloudless conditions. Then, we solve this
tomography problem, to recover the distribution. Recovery is
formulated as an optimization that minimizes a cost function. We
derive the gradient of this cost function, to enable efficient
optimization.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \vspace{-0.25cm}
\section{Theoretical Background}
\label{sec:back}
\vspace{-0.2cm}

\noindent {\bf Extinction}: Sun rays (SR) irradiate a small volume
that includes particles of a certain type.  Each particle has an {\em
  extinction cross section} $\sigma$ (units ${\rm m}^2$) for
interacting with the irradiance. The volume includes many particles of
this type. The denser the volume, the stronger is the extinction.  The
density of the particles is $n$ (units $1/{\rm m}^3$). Per unit
volume, the {\em extinction coefficient} $\beta$ (units $1/{\rm m}$)
is
\begin{align}
  \beta= \sigma n \;.
  \label{eq:extinctc}
\end{align}
The volume has infinitesimal length $dl$. Then, the relative portion
of extinct SR irradiance is the unitless differential {\em optical
  depth}
\begin{align}
  d\tau= \beta dl=\sigma n dl \;.
  \label{eq:extinct}
\end{align}
The optical depth aggregates in extended propagation:
\begin{align}
  \tau= \int d\tau=\int \beta dl=\int \sigma n dl \;.
  \label{eq:tau}
\end{align}
Through an attenuating volume, the {\em transmittance} exponentially
decays with the optical depth:
\begin{align}
  t=\exp^{-\tau}
  \label{eq:beer-lambert}
\end{align}

\noindent {\bf Scattering}: Interaction of a single particle with the
irradiance is by absorption and scattering. The weight of scattering
(to all directions), relative to the total extinction is given by the
unitless {\em single scattering albedo} $\varpi$ of the particle. The
{\em scattering coefficient} $\alpha$ (units $1/{\rm m}$) is
\begin{align}
  \alpha= \varpi\beta=\varpi \sigma n \;.
  \label{eq:alph}
\end{align}
Scattering is generally anisotropic. Its angular distribution is
determined by a {\em phase function} $P$. Part of the light scatters
towards a camera's line of sight (LOS), as illustrated in
Fig.~\ref{fig:groundgrid}. The angle between the SR and LOS is the
scattering angle $\Phi^{\rm scatter}$. The phase function $P(\Phi^{\rm
  scatter})$ is normalized: its integral over all solid angles is
unit. The {\em angular scattering coefficient} $\tilde\alpha$ (units
$1/({\rm m}\cdot{{\rm strad}})$) is
\begin{align}
  \tilde\alpha(\Phi^{\rm scatter}) = \varpi\beta P(\Phi^{\rm scatter})
  = \varpi \sigma n P(\Phi^{\rm scatter}).
  \label{eq:alphabasic}
\end{align}

The phase function is often approximated by a parametric expression,
as the Henyey-Greenstein function,
\begin{align}
  P_g(\Phi^{\rm scatter})\approx \frac{3} {8\pi} \frac{(1 -
    g^2)(1+\mu^2)} {(2+g^2)(1 + g^2 - 2g\mu)^\frac{3}{2}}
  \label{eq:aerosol_scatter}
\end{align}
where $g$ is an {\em anisotropy parameter} and
\begin{align}
  \mu\equiv \cos \Phi^{\rm scatter}.
  \label{eq:mu}
\end{align}

\noindent {\bf Air molecules}: Scattering by air molecules follows the
{\em Rayleigh} model: for wavelength $\lambda$ the phase function is
\begin{align}
  P^{\rm air}\left[\mu(k)\right] &= \frac{3}{16\pi}(1+\mu^2)
  \label{eq:rayleighP}
\end{align}
and $\varpi^{\rm air}$=1. Air molecular density $n^{\rm air}$ falls
off approximately exponentially with altitude, with a
characteristic~\cite{Levi1980} falloff height $H^\mathrm{air}=8\
km$. Consequently~\cite{Levi1980}, the coefficients for extinction and
scattering by air molecules are modeled by
\begin{align}
  \alpha^{\rm air}(h, \lambda)=\beta^{\rm air}(h, \lambda) \approxeq
  \frac{1.09 \times 10^{-3}}{\lambda^4}
  \exp(-h/H^\mathrm{air}) %\left[-\frac{h}{H^\mathrm{air}}\right].
  \label{eq:rayleighbeta}
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \vspace{-0.25cm}
\section{Image Acquisition Principle}
\label{sec:setup} \vspace{-0.2cm}

The atmosphere has a non-uniform distribution of scattering
aerosols. We seek to recover this 3D volumetric distribution. To solve
this problem, we formulate a process analogous to tomographic
reconstruction. The scene is projected to multiple directions. Thus, a
system captures integrated radiance as a function of direction and 2D
projected location. Any voxel in the atmosphere contributes to the
integrated radiance in multiple directional measurements
(Fig.~\ref{fig:groundgrid}). The measurements then provide a large
number of constraints, to recover the distribution: the density and
type of aerosols, per voxel.
\begin{figure*}[t!]
  \begin{center}
    \vspace{-0.4cm}
    \includegraphics[width=0.99\linewidth]{images/groundtomog24.eps}
  \end{center}\vspace{-1.5cm}
  \caption{\small Lightfield imaging through a volumetric distribution
    in the atmosphere, using ground-based cameras. Our tomographic
    reconstruction follows the acquisition. The atmosphere is analyzed
    in a voxel grid.}
  \label{fig:groundgrid}
\end{figure*}

Earth's radiation is sampled both spatially and angularly from
space~\cite{breon,diner,matronchik,vanMol}. However, current multiview
spaceborne systems are very coarse. For example, aerosol retrieval in
the most complex instrument, MISR, requires spatial blocks that are
about $17\times 17$km wide, each observed from just 9 views
along-track. This practically leads to vertical radiative transfer
analysis, degenerating atmospheric analysis to 1D. Experimental
airborne instruments under development~\cite{baxter,dinerDavis10} may
provide better, if still coarse resolution once they are operational.

We believe that high resolution atmospheric recovery can be achieved
by a large set of ground-based wide-angle~\cite{oliver, Peleg, kopal}
sky-cameras, looking upwards and capturing the sky lightfield.
Moreover, satellites image close to the orbit track (a 1D array of
viewpoints). In contrast, land-based cameras can form a 2D set of
viewpoints~\cite{Snavely} as easy or hard as a 1D
grid~\cite{Joshi}. However, before designing and setting up such a
very large scale system requires assessment of many details that had
not been explored. These include camera specifications needed to
recover typical distributions. Hence, a major theoretic and numerical
research phase is ahead.

This paper takes the first step, of formulating this principle and
demonstrating feasibility using realistic simulations. The formulation
of this paper can also apply to multiview images that are captured
from satellites, for future spaceborne instruments of this capacity.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \vspace{-0.25cm}
\section{Forward Model}
\label{sec:skymodel} \vspace{-0.2cm}

This section describes models for image formation. In the consecutive
section, we describe how the image formation is inverted to recover
the 3D sky structure.  Consider Fig.~\ref{fig:groundgrid}.
Ground-based wide-angle sky-cameras capture the sky lightfield from
below. Per viewpoint, the view azimuth and elevation angles (relative
to the zenith and north) are encapsulated in the vector
${\bf{\Theta}}$.  A raw image is $i_{\bf x}({\bf{\Theta}})$: a
camera's location ${\bf x}$ is fixed, while any pixel in a frame
represents an angle ${\bf{\Theta}}$.  The measured radiance in all
locations ${\bf x}$ and angles ${\bf{\Theta}}$ is the {\em lightfield}
$I({\bf x},{\bf{\Theta}})$. Individual images are slices of the
lightfield.
% Alternatively, several ground-based cameras can be mounted on moving
% vehicles, assuming that the aerosol distribution does not change
% much during the time-lapse of the camera trajectory.

Sunlight is scattered in the medium, creating {\em
  airlight}~\cite{fattal,narasimhan2,Hschechner2,tan}. For each view
${\bf{\Theta}}$ and ground location ${\bf x}$, the airlight is
generally different, depending on the distribution of aerosols both
along a LOS, and along the SRs that hit the LOS.  As
Fig.~\ref{fig:groundgrid} shows, different LOSs (distinct ${\bf
  x},{\bf{\Theta}}$) cut through different sets atmospheric
voxels. Thus, for each view, the radiance scattered from a voxel is
attenuated differently, depending on the distribution of aerosols
along the LOS~\cite{hansen}.

A forward model takes as input the aerosol distribution, and outputs images as if captured at various viewpoints. The model calculates radiative transfer, based on fixed parameters as the Sun's characteristics. We use a single scattering model, in which any SR changes direction (scatters) only once on the way to the camera. This model is valid in atmospheres that are not very dense: inside fog and clouds this model does not apply. Under single-scattering approximation, radiative transfer has three steps (Fig.~\ref{fig:groundgrid}):\\

\vspace{-0.35cm}
\noindent
1. Attenuation of a SR propagating from the top of the atmosphere
(TOA) to voxel $k$. \\

\vspace{-0.35cm}
\noindent
2. Light scattering at voxel $k$, towards a camera.\\

\vspace{-0.35cm}
\noindent
3. Light attenuation on the LOS from voxel $k$ to the camera.

\noindent
Steps 1 and 3 involve the optical depth along the light path to and
from voxel $k$, analyzed in Sec.~\ref{sec:optical-depth}.  Step 2
involves the scattering coefficient at voxel $k$, described in
Sec.~\ref{sec:scattering}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Optical Depth}
\label{sec:optical-depth}

An infinitesimal volume contains aerosols. The extinction and
scattering coefficients due to aerosols in this volume are,
respectively
\begin{equation}
  \beta^{\rm aerosol}=\sigma^{\rm aerosol} n^{\rm aerosol}
  \;\;,\;\;
  \alpha^{\rm aerosol}=\varpi^{\rm aerosol} \beta^{\rm aerosol}
  \;.
  \label{eq:betaer}
\end{equation}
In addition, the infinitesimal volume contain also air
molecules. Hence, the overall extinction and scattering coefficients
in this small volume are, respectively
\begin{equation}
  \beta=\beta^{\rm air}+\beta^{\rm aerosol}
  \;\;,\;\;
  \alpha=\alpha^{\rm air}+\alpha^{\rm aerosol}
  \;.
  \label{eq:betair}
\end{equation}

Any voxel generally contains both air molecules and aerosols. Voxels
are indexed by $k$ or $q$. As approximation, assume that within any
voxel, the molecular parameters $\{ \beta^{\rm air}(k),\alpha^{\rm
  air}(k)\}$ and the aerosol parameters $\{ \sigma^{\rm
  aerosol}(k),\varpi^{\rm aerosol}(k), n^{\rm aerosol}(k),g(k)\}$ are
constants, e.g., corresponding to the value at each voxel
center. These parameters may change between voxels.

A voxel's vertical geometric thickness is $\Delta z$.  The Sun zenith
angle is $\Phi^{\rm SR}$. Denote a SR line segment between the TOA and
the center of voxel $k$ by $[{\rm SR},k]$. This SR intersects voxel
$q$. The geometric length of this intersection line segment is $\Delta
z f_{\rm SR}(q)/\cos\Phi^{\rm SR}$: the factor \mbox{$f_{\rm SR}(q)\in
  [0,1]$} expresses how much of the vertical thickness of $q$ is
crossed by the SR. Following Eqs.~(\ref{eq:tau},\ref{eq:betair}), the
optical depth between the TOA and voxel $k$ is
\begin{equation}
  \tau_{\rm SR}(k)=\frac{1}{\cos\Phi^{\rm SR}}
  \sum_{q \in[{\rm SR},k]}
  \beta(q)\Delta z  f_{\rm SR}(q)
  \;\;.
  \label{eq:lSRn}
\end{equation}
Based on Eqs.~(\ref{eq:betaer},\ref{eq:betair},\ref{eq:lSRn}),
\begin{equation}
  \tau_{\rm SR}(k)=
  \tau_{\rm SR}^{\rm air}(k) +  \tau_{\rm SR}^{\rm aerosol}(k)
  \;\;,
  \label{eq:lSRaa}
\end{equation}
where
\begin{equation}
  \tau_{\rm SR}^{\rm air}(k)=\frac{\Delta z}{\cos\Phi^{\rm SR}}
  \sum_{q \in[{\rm SR},k]}
  \beta^{\rm air}(q)  f_{\rm SR}(q)
  \label{eq:lSRair}
\end{equation}
and
\begin{equation}
  \tau_{\rm SR}^{\rm aerosol}(k)=\frac{\Delta z}{\cos\Phi^{\rm SR}}
  \sum_{q \in[{\rm SR},k]}
  \sigma^{\rm aerosol}(q) n^{\rm aerosol}(q) f_{\rm SR}(q)
  \label{eq:lSRaeros}
\end{equation}

The number of voxels is $N_{\rm voxels}$. All their associated
variables are concatenated into columns stack vectors, each $N_{\rm
  voxels}$ long. Aerosol densities and cross sections in all voxels
are stored in the respective vectors ${\bf n}$ and
$\vect{\sigma}$. Define a $N_{\rm voxels}\times N_{\rm voxels}$ sparse
matrix ${\bf D}^{\mathrm{Sun \rightarrow voxel}}$, whose element
$(k,q)$ is
\begin{equation}
  D^{\mathrm{Sun \rightarrow voxel}}(k,q) =
  \left\{
    \begin{array}{ll}
      \frac{\Delta zf_{\rm SR}(q)}{\cos\Phi^{\rm SR}}
      & \mbox{ if $q \in[{\rm SR},k]$} \\
      0  & \mbox{ otherwise}
      \label{eq:Dsv}
    \end{array}
  \right.
  .
  \hspace{-0.05cm}
\end{equation}
Eqs.~(\ref{eq:lSRaeros},\ref{eq:Dsv}) lead to a vector of SR optical
depths due to aerosols (See Fig.~\ref{fig:optidepth}):
\begin{equation}
  \vect{\tau}_{\rm SR}^{\rm aerosol}=
  {\bf D}^{\mathrm{Sun \rightarrow voxel}}
  (\vect{\sigma} \odot {\bf n})
  \label{eq:tauSRvec1}
\end{equation}
where the operator $\odot$ denotes the Hadamard (element-wise)
product.
\begin{figure}
  \centering
  \yoavcomment{\def\svgwidth{\columnwidth}\input{images/optical_distance.pdf_tex}}
  \caption{\small Irradiance of voxel $k$ is attenuated by voxels
    above it on a SR. The scattered radiance from voxel $k$ is then
    attenuated by voxels below it on a LOS to camera $c$. The overall
    attenuation of any voxel $k$ by all voxels $q$ is equivalent to an
    optical depth affecting $k$. It is formalized in matrix-vector
    operations. }
  \label{fig:optidepth}
\end{figure}
Similarly, Eq.~(\ref{eq:lSRair}) leads to a vector of SR optical
depths due to air, $\vect{\tau}_{\rm SR}^{\rm air}$.

Analogously, for camera $c$, denote by $[{\rm LOS}_c,k]$ a LOS bounded
between the camera and the center of voxel $k$. The LOS zenith angle
is $\Phi^{\rm LOS}_c$.  This LOS intersects voxel $q$. The geometric
length of this intersection line segment is $\Delta z f_{\rm
  LOS_c}(q)/\cos\Phi^{\rm LOS}_c$: the factor $f_{\rm LOS_c}(q)\in
[0,1]$ expresses how much of the vertical thickness of $q$ is crossed
by the LOS.  Analogously to
Eqs.~(\ref{eq:lSRaa},\ref{eq:lSRair},\ref{eq:lSRaeros}), The optical
depth %path length
along a LOS to $k$ is
\begin{equation}
  \tau_{{\rm LOS}_c}(k)=
  \tau_{{\rm LOS}_c}^{\rm air}(k) +  \tau_{{\rm LOS}_c}^{\rm aerosol}(k)
  \;\;,
  \label{eq:lLOSaerot}
\end{equation}
where
\begin{equation}
  \tau_{{\rm LOS}_c}^{\rm air}(k)=\frac{\Delta z}{\cos\Phi^{\rm LOS}_c}
  \sum_{q \in[{\rm LOS}_c,k]}
  \beta^{\rm air}(q)  f_{\rm LOS_c}(q)
  \label{eq:lLOSair}
\end{equation}
and
\begin{align}
  \label{eq:lLOSaeros}
  \tau_{{\rm LOS}_c}^{\rm aerosol}(k)=
  ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ ~~~~~~~~~~~~~~~~~~%\nonumber
  ~~~~~
  \\
  % ~~~~
  \frac{\Delta z}{\cos\Phi^{\rm LOS}_c} \sum_{q \in[{\rm LOS}_c,k]}
  \sigma^{\rm aerosol}(q) n^{\rm aerosol}(q) f_{\rm LOS_c}(q).
  \nonumber
\end{align}
Eq.~(\ref{eq:lLOSaeros}) leads to a vector of LOS optical depths due
to aerosols:
\begin{equation}
  \vect{\tau}_{{\rm LOS}_c}^{\rm aerosol}=
  {\bf D}^{\mathrm{voxel \rightarrow cam}}_c
  (\vect{\sigma} \odot {\bf n})
  \label{eq:tauLOSvec1}
\end{equation}
Here ${\bf D}^{\mathrm{voxel \rightarrow cam}}_c$ is a $N_{\rm
  voxels}\times N_{\rm voxels}$ sparse matrix whose element $(k,q)$ is
\begin{equation}
  D^{\mathrm{voxel \rightarrow cam}}_c(k,q) =
  \left\{
    \begin{array}{ll}
      \frac{\Delta z f_{\rm LOS_c}(q)}{\cos\Phi^{\rm LOS}_c}
      & \mbox{ if $q \in[{\rm LOS}_c,k]$} \\
      0  & \mbox{ otherwise}
      \label{eq:Dvc}
    \end{array}
  \right.
  .
  \hspace{-0.05cm}
\end{equation}
Similarly, Eq.~(\ref{eq:lSRair}) leads to a vector of LOS optical
depths due to air, $\vect{\tau}_{{\rm LOS}_c}^{\rm air}$.

The fixed sparse matrices ${\bf D}^{\mathrm{Sun \rightarrow voxel}}$
and ${\bf D}^{\mathrm{voxel \rightarrow cam}}_c$ are compounded to a
single matrix
\begin{equation}
  {\bf D}_c=
  {\bf D}^{\mathrm{Sun \rightarrow voxel}}+
  {\bf D}^{\mathrm{voxel \rightarrow cam}}_c
  \;\;.
  \label{eq:Dtotal}
\end{equation}
Similarly, the molecular optical depths along all trajectories between
the sun, each voxel and camera $c$ are compounded to a vector
\begin{equation}
  \vect{\tau}_c^{\rm air}=
  \vect{\tau}_{\rm SR}^{\rm air}+
  \vect{\tau}_{{\rm LOS}_c}^{\rm air}
  \;\;.
  \label{eq:tauairtotal}
\end{equation}
Using Eq.~(\ref{eq:rayleighbeta}) for $\beta^{\rm air}$ in all voxels,
the vector $\vect{\tau}_c^{\rm air}$ is fixed. Thus
$\vect{\tau}_c^{\rm air}$ can be calculated once, per sun zenith
angle.

Using
Eqs.~(\ref{eq:lSRaa},\ref{eq:tauSRvec1},\ref{eq:lLOSair},\ref{eq:tauLOSvec1},\ref{eq:Dtotal},\ref{eq:tauairtotal}),
the total optical depths corresponding to LOSs (of camera $c$) and SRs
that cross all voxels comprise the vector
\begin{equation}
  \vect{\tau}_c= \vect{\tau}_c^{\rm air}
  + {\bf D}_c (\vect{\sigma} \odot {\bf n})
  \;\;.
  \label{eq:tautotal}
\end{equation}
This vector depends on the variables $\vect{\sigma}$ and ${\bf n}$
(See Fig.~\ref{fig:optidepth}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Scattering}
\label{sec:scattering}

Per camera $c$ and voxel $k$, the lines $[{\rm SR},k]$ and $[{\rm
  LOS}_c,k]$ intersect at a fixed angle $\Phi^{\rm scatter}_c(k)$,
which is pre-calculated.
% Eq.~(\ref{eq:mu}) yields its cosine, $\mu_c(k)$.
Column-stacking all voxels yields the vector representation of all
scattering angles in the domain, $\vect{\Phi}^{\rm scatter}_c$ per
camera $c$. Similarly, the single-scattering albdeos $\varpi^{\rm
  aerosol}$ in all voxels are stacked to vector $\vect{\varpi}$.
Using
Eqs.~(\ref{eq:alphabasic},\ref{eq:rayleighP},\ref{eq:betaer},\ref{eq:betair}),
the angular scattering coefficients across the domain are expressed in
vector form by
\begin{align}
  \tilde{\vect{\alpha}}^{\rm air}_c &=
  \vect{\beta}^{\rm air} \odot P^{\rm air}(\vect{\Phi}^{\rm scatter}_c)   \nonumber \\
  \tilde{\vect{\alpha}}^{\rm aerosol}_c &= \vect{\varpi}\odot
  \vect{\sigma} \odot {\bf n} \odot P^{\rm aerosol}_{\bf
    g}(\vect{\Phi}^{\rm scatter}_c)
  \label{eq:alpha_matrix}
\end{align}
The vector ${\bf g}$ column-stacks the anisotropy parameter of the
phase function, in all voxels. We use this notation, to indicate a
parametric phase function as in Eq.~(\ref{eq:aerosol_scatter}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Image Capture}
\label{sec:captured-image}

Compounding the attenuation of irradiance along both $[{\rm SR},k]$
and $[{\rm LOS}_c,k]$, and scattering by voxel $k$ towards the camera
(Eqs.~\ref{eq:beer-lambert},\ref{eq:betair},\ref{eq:tautotal}), the
radiant power contributed by the voxel is
\begin{equation}
  p_c(k)= L^{\rm TOA}
  [\tilde{\alpha}^{\rm air}_c(k) + \tilde{\alpha}^{\rm aerosol}_c(k)]
  e^{-[\tau_c(k)]}
  \;\;.
  \label{eq:ick}
\end{equation}
A column-stack vector of all voxel contributions is
\begin{equation}
  {\bf p}_c= L^{\rm TOA}
  [\tilde{\vect{\alpha}}^{\rm air}_c + \tilde{\vect{\alpha}}^{\rm aerosol}_c]
  \odot e^{-\vect{\tau}_c}
  \;\;.
  \label{eq:pc}
\end{equation}

A camera pixel collects light from a narrow cone in the
atmosphere. The cone includes some voxels, intersects with some other
voxels, and oblivious to all the rest.  Thus light power captured at a
pixel is weighted sum of the power $p_c(k)$ from all voxels: most
voxels have null weight, since they fall outside the pixels's conic
field of view, while the rest of the voxels are aggregated.

This linear sum is formulated by a matrix operation over ${\bf
  p}_c$. The measured image is column-stacked to a vector $N_{\rm
  pix}$ long. It is given by
\begin{equation}
  {\bf i}_c= {\vect{\Pi}}_c{\bf p}_c
  \;\;.
  \label{eq:IcP}
\end{equation}
The $N_{\rm pix}\times N_{\rm voxels}$ sparse matrix ${\vect{\Pi}}_c$
expresses {\em projection} of light into camera $c$. For pixel
$\Theta$ and voxel $k$, there is an {\em intersection} volume ${\bf
  V}_c(\Theta,k)$ between the cone volume of the pixel's field of
view, and the voxel, as illustrated in Fig.~\ref{fig:projection}.
\begin{figure}
  \centering
  \yoavcomment{\def\svgwidth{\columnwidth}\input{images/sensor.pdf_tex}}
  \caption{\small An image pixel captures light emanating in a cone in
    space. Voxels that fully or partially fall in this cone are sensed
    by the pixel. Factoring out scattering and attenuation (which are
    dealt elsewhere in our analysis), the relative contribution of
    each voxel depends on its distance and intersection with the
    cone. This is formalized a matrix-vector operation.}
  \label{fig:projection}
\end{figure}
The corresponding element in the matrix ${\vect{\Pi}}_c$ is
\begin{equation}
  \Pi_c(\Theta,k)=
  % \frac{a}{|{\bf V}(k)|}
  a
  \iiint_{{\bf V}_c(\Theta,k)}
  \frac{1}{[R({\bf v})]^2} d{\bf v}
  \;\;,
  \label{eq:Pc}
\end{equation}
where $a$ is the effective area of the camera lens.
% and $|{\bf V}(k)|$ is the volume of voxel $k$.
The distance $R({\bf v})$ is measured between the camera and an
infinitesimal volume element at ${\bf v}\in {\bf V}_c(\Theta,k)$.

Combining Eqs.~(\ref{eq:tautotal},\ref{eq:pc},\ref{eq:IcP}), the
captured image is thus
\begin{align}
  {\bf i}_c= L^{\rm TOA} {\vect{\Pi}}_c \left\{
    (\tilde{\vect{\alpha}}^{\rm air}_c + \tilde{\vect{\alpha}}^{\rm
      aerosol}_c) \odot e^{-[\vect{\tau}_c^{\rm air} + {\bf D}_c
      (\vect{\sigma} \odot {\bf n})]} \right\}
  \label{eq:bigIc}
\end{align}
Most often, there is only a single type of aerosol over a
scene. Hence, the three-element vector $[{\sigma}^{\rm
  aerosol},\varpi^{\rm aerosol},g]$ is uniform across the scene, but
the density distribution ${\bf n}$ is variable. Then,
Eq.~(\ref{eq:bigIc}) degenerates to
\begin{align}
  {\bf i}_c= L^{\rm TOA} {\vect{\Pi}}_c \Big\{ e^{-(\vect{\tau}_c^{\rm
      air} + {\sigma}^{\rm aerosol} {\bf D}_c {\bf n}) }\odot
  \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;
  \nonumber \\
  [\tilde{\vect{\alpha}}^{\rm air}_c + %\right.
  \varpi^{\rm aerosol}\sigma^{\rm aerosol} P^{\rm
    aerosol}_g(\vect{\Phi}^{\rm scatter}_c) \odot{\bf n} ] \Big\} .
  \label{eq:bigIA}
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Inverse Problem}
\label{sec:inverse-problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Unknowns and Prior Knowledge}
\label{sec:prior}

In general terms, the variables are the set
\begin{equation}
  {\cal V}\equiv
  \left\{
    {\bf n},
    \left\{
      \vect{\varpi}_{\lambda},\vect{\sigma}_{\lambda},{\bf g}_{\lambda}
    \right\}_{\lambda}
  \right\}
  \;\;,
  \label{eq:calV}
\end{equation}
where $\lambda$ denotes the spectral (color) channel.  There are
$N_{\rm spectral}$ spectral channels.  Thus, the number of unknown
variables is in principle \mbox{$|{\cal V}|=N_{\rm voxels}(1+3N_{\rm
    spectral})$}.

Prior knowledge reduces the unknowns. Sec.~\ref{sec:captured-image}
mentions a prior: in many cases a single aerosol type dominates a
scene. Then, the three-element vector $[\varpi^{\rm
  aerosol},{\sigma}^{\rm aerosol},g]$ is rather spatially uniform. In
such cases, the set of unknowns reduces to $ {\cal V}= \left\{ {\bf
    n}, \left\{ {\varpi}^{\rm aerosol}_{\lambda},{\sigma}^{\rm
      aerosol}_{\lambda},g_{\lambda} \right\}_{\lambda} \right\} $ and
\mbox{$|{\cal V}|=N_{\rm voxels}+3N_{\rm spectral}$}.

Furthermore, the elements in the aerosol characteristics set $\left\{
  {\varpi}^{\rm aerosol}_{\lambda},{\sigma}^{\rm
    aerosol}_{\lambda},g_{\lambda} \right\}_{\lambda}$ is not
arbitrary, since all these characteristics have physical origins. This
set is constrained to comply with what nature allows (aerosol
classes).  This significantly constrains the unknown
characteristics. State-of-the-art aerosol retrieval, as used by NASA's
MISR products indeed employ an aerosol-class prior. Types of aerosol
mixtures are discretized to $N_{\rm models}=74$
classes~\cite{matronchik}, called {\em models}. Each class $m$
represents aerosols having a distinct characteristics set.  State of
the art retrieval is 1D (only vertical radiative transfer per large
areas). Nevertheless, we use this prior in our 3D problem. Then, the
set of unknowns reduces to $ {\cal V}= \left\{ {\bf n},m \right\}$ and
\mbox{$|{\cal V}|=N_{\rm voxels}+N_{\rm models}$}.

The atmosphere is photographed from $N_{\rm views}$ viewpoints. Each
photograph provides $N_{\rm pix}$ measurements (pixels), {\em
  including} their possible division to color channels. Hence, the
data at our disposal comprises $N_{\rm views}N_{\rm pix}$
measurements.  For the recovery task to be well-posed, it is necessary
that $N_{\rm views}N_{\rm pix}>|{\cal V}|$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Objective Function}
\label{sec:objective-function}

The inverse problem (tomography) seeks recovery of the 3D aerosol
distribution in the atmosphere. From
Eqs.~(\ref{eq:alpha_matrix},\ref{eq:bigIc},\ref{eq:bigIA},\ref{eq:calV}),
the images $\{ {\bf i}_c\}_{c=1}^{N_{\rm views}}$ depend on the
variables set ${\cal V}$, particularly on the unknown field ${\bf
  n}$. The recovery task is phrased as optimization of an objective
function, that fits the image-formation
model~(\ref{eq:bigIc},\ref{eq:bigIA}) to measured photographs $\{ {\bf
  i}^{\rm measured}_c\}_{c=1}^{N_{\rm views}}$:
\begin{equation}
  \label{eq:mingenobjective}
  \hat{\cal V} = \argmin_{{\cal V}\in {\cal C}} F({\cal V})
\end{equation}
where
\begin{equation}
  \label{eq:genobjective}
  F({\cal V}) = \sum_{c=1}^{N_{\rm views}}
  \| {\bf i}^{\rm measured}_c - {\bf i}_c({\cal V})\| + {\rm Regularization}({\cal V})
  \;.
\end{equation}
Here ${\cal C}$ is a constraints set, which encapsulates part of the
prior knowledge, as well as nonnegativity bounds, while $\|~\|$ is
some error-distance measure.  Here ${\rm Regularization}({\cal V})$
describes spatial regularization terms of the variables, such as
smoothness, or a blobby or layered structure.

Spatial regularization can help overcome lack of data, if $N_{\rm
  views}$ is too small, and it is definitely helpful. In this work,
however, we formulate and demonstrate the basic principle of our
tomography task, hence we proceed without regularization. Relying on
prior knowledge described in Sec.~\ref{sec:prior}, the problem
complexity is by far dominated by the $N_{\rm voxels}$-large array
${\bf n}$, which is continuously valued. Hence, we focus here on
retrieving ${\bf n}$ through continuous optimization. Relying on the
least-squares measure, we seek:
\begin{equation}
  \label{eq:minobjectiveA}
  \hat{\bf n} =
  \argmin_{{\bf n}} F({\bf n})
\end{equation}
where
\begin{equation}
  \label{eq:objectiveA}
  F({\bf n})
  = \sum_{c=1}^{N_{\rm views}}
  \| {\bf i}^{\rm measured}_c - {\bf i}_c({\bf n})\|^2_2
\end{equation}
This is solved using standard optimization tools. Fast optimization
that leads to to local minima ($F$ is not convex) exploits the
gradient of $F$.

Since our model uses close-form expressions, $F$ is
differentiable. The gradient of $F$ with respect to ${\bf n}$ is
\begin{align}
  \Grad{{\bf n}} F = -2\sum_{c=1}^{N_{\rm views}}
  \transpose{\left[{\bf J}_{{\bf i}_c}({\bf n})\right]} [{\bf i}^{\rm
    measured}_c - {\bf i}_c({\bf n})] \;.
  \label{eq:gradient1}
\end{align}
Here the matrix ${\bf J}_{{\bf i}_c}({\bf n})$ is the Jacobian of the
vector ${\bf i}_c$ with respect to ${\bf n}$. Element $(\Theta,k)$ of
this matrix differentiates the intensity in pixel ${\bf \Theta}$ (in
viewpoint $c$) with variation of the density at voxel $k$, i.e.,
$\partial i_c({\bf \Theta})/\partial{n^{\rm aerosol}(k)}$.

Denote by $\OpDiag{\vect{v}}$ conversion of vector $\vect{v}$ into a
diagonal matrix, whose main diagonal elements correspond to the
elements of $\vect{v}$.  A close-form expression of our Jacobian is
\begin{align}
  {\bf J}_{{\bf i}_c}({\bf n}) &= L^{\rm TOA}({\bf A}-{\bf B}){\bf C}
  ,
  \label{eq:gradient2}
\end{align}
where
\begin{align}
  \label{eq:gradient3}
  {\bf A} &= \varpi^{\rm aerosol}\sigma^{\rm aerosol}
  \OpDiag{ P_g^{\rm aerosol}(\vect{\Phi}^{\rm scatter}_c)} \nonumber\\
  {\bf B} &= {\sigma}^{\rm aerosol}\transpose{{\bf D}_c} \nonumber\\
  & \;\;\; \OpDiag{[\tilde{\vect{\alpha}}^{\rm air}_c + \varpi^{\rm
      aerosol}\sigma^{\rm aerosol} P^{\rm aerosol}_g(\vect{\Phi}^{\rm
      scatter}_c) \odot{\bf n}
    ]} \nonumber \\
  {\bf C} &= \OpDiag{\exp[-(\vect{\tau}_c^{\rm air} + {\sigma}^{\rm
      aerosol} {\bf D}_c {\bf n})]} \transpose{{\vect{\Pi}}_c}
\end{align}
For a detailed derivation of
Eqs.~(\ref{eq:gradient2},\ref{eq:gradient3}), we refer the reader to
the supplementary material.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulations}
\label{sec:simul}

To test the model, we made simulations based on atmospheric models in the literature. Here are the details.\\

\noindent{\bf Sun}: The sun is at zenith angle $\Phi^{\rm
  SR}=45^o$. Its $L^{\rm TOA}$ is obtained
from~\cite{BBradiance,sun_composition}, with respective red-green-blue
ratios of
$L^{\rm TOA}_{\rm R}:L^{\rm TOA}_{\rm G}:L^{\rm TOA}_{\rm B}=255:236:224$.\\

\noindent{\bf Dimensions}: The domain of sky we simulate and recover has area of $50\times 50$km, and it extends from the ground up to altitude of 10km. This volume is divided to a $50\times50\times100$ grid, where each voxel has area of $1\times1$km and vertical thickness of 100 meters. The dimension of the problem is thus $250,000$ variables.\\

\noindent{\bf Aerosol}: We used particle-type 6 from the aerosol list
in Ref.~\cite{Martonchik2009}. This is a spherical non-absorbing
sea-salt/organic particle, whose anisotropy parameter per color
channel is $[g_{\rm R},g_{\rm G},g_{\rm B}]=[0.763,0.775,0.786]$. Its
corresponding extinction cross sections are $[\sigma^{\rm
  aerosol}_{\rm R}, \sigma^{\rm aerosol}_{\rm G}, \sigma^{\rm
  aerosol}_{\rm B}]=[16.5,16.2,15.9][1/\mu m^2]$.  Its
single-scattering albedo is $\varpi^{\rm aerosol}=1$ at all channels.

We simulated various spatial distributions. All were set as a product
of two functions. The first function is an exponential
decay~\cite{Levi1980} with altitude $h(k)$ of voxel $k$,
\begin{align}
  w_1[h(k)] = \frac{1} {\sigma^{\rm aerosol}_{\rm G} \mathrm{MFP^{sea
        level}}} \exp\left[-\frac{h(k)}{H^\mathrm{aerosol}}\right],
\end{align}
where $H^\mathrm{aerosol}=2\ km$. This expresses a general trend of
reduced atmospheric density with height. Here the mean free path (MFP)
of the aerosols layer at sea level was set to $\mathrm{MFP^{sea
    level}}=1/5~{\rm[1/km]}$.
% is a typical characteristic thickness of the aerosol
% layer~\cite{Levi1980}.

The second function, $w_2$, expresses a clustered nature of aerosol
``clouds''. We define blobs in the form of 3D ellipsoids. There may be
multiple ellipsoids suspended.  Then
\begin{equation}
  w_2(k) =
  \left\{
    \begin{array}{ll}
      1  & \mbox{ if $k\in$ any blob cluster} \\
      0  & \mbox{ otherwise}
      \label{eq:f2}
    \end{array}
  \right.
  .
  \hspace{-0.05cm}
\end{equation}
The true aerosol distribution is then
\begin{equation}
  \label{eq:ntrue}
  {\bf n}^{\rm true}={\bf w}_1\odot {\bf w}_2  .
  \hspace{-0.05cm}
\end{equation}
As an example, we created a {\em Haze Blobs} scene
(Fig.~\ref{fig:simulation1}) using two ellipsoid blobs: \begin{figure}
  \centering
  \yoavcomment{\includegraphics[width=\columnwidth]{images/simulation1}}
  \caption{\small Simulated 3D aerosol distribution composed of {\em
      Haze Blobs}. Its reconstruction is shown in
    Fig.~\ref{fig:front}. Color represents aerosol density. The
    density units are $10^{15}~{\rm particles}/{\rm cm}^3.$}
  \label{fig:simulation1}
\end{figure}
the first is centered at altitude $3.3km$, and has horizontal width of $32km$ and vertical thickness of $2.8km$. The second ellipsoid is centered at altitude $6.6km$, and has lateral width of $24km$ and vertical thickness of $2.1km$. Note that the horizontal widths are much larger than the vertical ones. This is consistent with scales of atmospheric aerosol distributions.\\

\noindent{\bf Cameras}: Each ground-based camera has a hemispherical
field of view, capturing the entire sky from its viewpoint. We used 95
viewpoints on a grid, each view $\approx 4.4$km from its neighbors.
The resolution of each camera is low: $128\times 128$ pixels,
expressing the fact that cloudless (yet hazy) sky images are rather
smooth, hence there is no need for high image resolution. Each pixel
has full-well depth of 40000 electrons, which corresponds to the
brightest pixel value. Each measurement has photon noise, which is
Poissonian distributed, according to each image-pixel value $i_c$ and
the full-well depth. For example, Fig.~\ref{fig:simulation-results1}
shows two modeled photographs
\begin{figure}
  \centering
  \yoavcomment{\includegraphics[width=\columnwidth]{images/ref_images.pdf}}
  \caption{\small Rendered photographs (contrast enhanced)
    corresponding to different viewpoints, of sky that includes {\em
      Haze Blobs}. A yellow dot marks the Sun location. Dashed white
    circles mark elevation angles.}
  \label{fig:simulation-results1}
\end{figure}
of the  {\em Haze Blobs} scene that corresponds to Fig.~\ref{fig:simulation1}.\\

\noindent{\bf Computer}: We had access to an SGI-TNN cluster computer
that has 88 computer nodes consisting of two 2.40 GHz six-Core Xeon
with 8GB
DDR3 memory per core. Out of it, we used 8 nodes, corresponding to 96 cores. Each core was dedicated to rendering a modeled image (in each interaction) as if captured by one of the 95 cameras. Each iteration (varying the model according to the gradient-based algorithm and varying all the rendered images) took $\approx 7$sec.\\

\noindent{\bf Optimization}: We used the L-BFGS-B
algorithm~\cite{BFGS}. The algorithm was initialized by ${\bf
  n}=0$. Satisfactory convergence occurred in several thousand
iterations, taking a couple of hours to complete in total. For
example, Fig.~\ref{fig:objective} shows
\begin{figure}[t]
  \centering
  \yoavcomment{\includegraphics[width=\columnwidth]{images/objective.pdf}}
  % \includegraphics[width=\columnwidth]{images/objective.pdf}
  \caption{\small The decay of $\log F({\bf n})$ with the iteration
    index.}
  \label{fig:objective}
\end{figure}
the progression of the objective as a function of the iteration
number, when recovering the the {\em Haze Blobs} scene.

% 26, 48, 1320, 590
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Results}
\label{sec:optimization-results}

An iso-surface visualization of the original {\em Haze Blobs}
distribution is shown in Fig.~\ref{fig:simulation1}, while the
recovered 3D distribution is visualized in Fig.~\ref{fig:front}. The
distributions appear consistent, though there are errors that stem
from the random noise in the images. To quantify the results, we use a
relative error measure based on the $\ell_1$ norm
\begin{equation}
  \label{eq:ndiff}
  \epsilon=\frac{\| \hat{\bf n}  - {\bf n}^{\rm true} \|_1}
  {\| {\bf n}^{\rm true} \|_1} ,
  \hspace{-0.05cm}
\end{equation}
It is equivalent to the aerosol mass that is over and under-estimated
in all voxels, relative to the total aerosol mass in the scene.  In
the {\em Haze Blobs} case, $\epsilon=0.202$.

We created a {\em Haze Front} scene. Fig.~\ref{fig:simulation2}a shows
an iso-surface visualization of this distribution.
\begin{figure}
  \centering
  \yoavcomment{\includegraphics[width=\columnwidth]{images/simulation2}}
  \caption{\small [a] Simulated 3D aerosol distribution composed of a
    {\em Haze Front}. [b] Its reconstruction. Color represents aerosol
    density. The density units are $10^{15}~{\rm particles}/{\rm
      cm}^3.$}
  \label{fig:simulation2}
\end{figure}
Here an ellipsoid degenerated to an elliptic cylinder that partly
enters the analyzed volume. It is centered at altitude $5km$, has
vertical thickness of $4km$ and length of $32km$ (only part of which
enters the volume). The haze front stretches across the width of the
volume.  Fig.~\ref{fig:simulation2}b visualizes the recovered 3D
distribution of this scene. For the {\em Haze Front},
$\epsilon=0.198$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:conclusions}

We describe a new way to sense the 3D atmosphere. We show feasibility
of departing from common 1D aerosol distribution models, in favor of
detailed 3D analysis.  The paper describes both a novel data
acquisition system concept for this recovery task, and a dedicated
algorithm for reconstructing aerosol distributions.

Even without regularization, the reconstruction algorithm handles
arbitrary density distributions. Using more prior knowledge,
e.g. spatial smoothness will improve the recovery.  The algorithm uses
a single scattering model. This approximation is valid when the
optical depth is small, which is common. To handle dense atmospheric
conditions (smoke, clouds), multi-scattering must be incorporated.

The goal, of coarse, is to deploy such a system on a very large
scale. However, this must first ensure that camera units are both cost
effective and field-robust, and that their positioning density is
proper. This will require a long-term engineering effort. Camera
specifications should be set to optimally recover a wide range of
aerosol distributions. This requires a major theoretic and numerical
research and development phase.  The new sensing modality and object
of interest offer a new domain for computational photography
problems. For example, how can compressed sensing principles be
employed here? How can priors be used to make data acquisition more
efficient? How can such a system be calibrated over time? Many such
questions and solutions would rise.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

{\small
  % \bibliographystyle{ieee} \bibliography{atmosphere}

  \begin{thebibliography}{10}\itemsep=-1pt

  \bibitem{baxter} B.~Baxter, B.~Hooper, J.~Williams, and J.~Dugan.
    \newblock Polarimetric remote sensing of ocean waves.  \newblock
    Proc. MTS/IEEE OCEANS 2009.

  \bibitem{bishop} T.~E. Bishop, S.~Zanetti, and P.~Favaro.  \newblock
    Light field superresolution.  Proc. IEEE ICCP 2009.

  \bibitem{breon} F.~Br\'{e}on.  \newblock An analytical model for the
    cloud-free atmosphere/ocean system reflectance.  \newblock {\em
      Remote Sensing of Environment}, 43:179 -- 192, 1993.

    % \bibitem{chang} G.~C. Chang and R.~W. Gould.  \newblock
    %   Comparisons of optical properties of the coastal ocean derived
    %   from satellite ocean color and in situ measurements.
    %   \newblock {\em Opt. Express}, 14:10149--10163, 2006.

  \bibitem{BBradiance} M.~Charity.  \newblock Blackbody color
    datafile, 2001.  \newblock [Online; accessed 6-August-2012].

    % \bibitem{chud} A.~Chudnovsky, A.~Kostinski, L.~Herrmann,
    %   I.~Koren, G.~Nutesku, and E.~Ben-Dor.  \newblock Hyperspectral
    %   spaceborne imaging of dust-laden flows: Anatomy of saharan
    %   dust storm from the Bod\'{e}l\'{e} depression.  \newblock {\em
    %   Remote Sensing of Environment}, 115:1013 -- 1024, 2011.

  \bibitem{oliver} O. Cossairt, D. Miau, and S.K. Nayar.  \newblock
    Gigapixel Computational Imaging.  \newblock Proc. IEEE ICCP 2011.

  \bibitem{matronchik} J.~M. David, D.~J. Diner, R.~Kahn,
    T.~P. Ackerman, M.~M. Verstraete, B.~Pinty, and H.~R. Gordon.
    \newblock Techniques for the retrieval of aerosol properties over
    land and ocean using multi-angle imaging.  \newblock {\em IEEE
      Trans. Geoscience and Remote Sensing}, 36:1212--1227, 1998.

    % \bibitem{martonchikBook} J.~M. David, D.~J. Diner, R.~Kahn,
    %   T.~P. Ackerman, M.~M. Verstraete, B.~Pinty, and H.~R. Gordon.
    %   \newblock Techniques for the retrieval of aerosol properties
    %   over land and ocean using multi-angle imaging.  \newblock {\em
    %   IEEE Trans. Geoscience and Remote Sensing}, 36:1212--1227,
    %   1998.

  \bibitem{dayan} U.~{Dayan}, B.~{Ziv}, T.~{Shoob}, and Y.~{Enzel}.
    \newblock {Suspended dust over southeastern Mediterranean and its
      relation to atmospheric circulations}.  \newblock {\em
      Int. J. Climatology}, 28:915--924, 2008.

  \bibitem{dinerDavis10} D.~J. Diner, A.~Davis, B.~Hancock, S.~Geier,
    B.~Rheingans, V.~Jovanovic, M.~Bull, D.~M. Rider, R.~A. Chipman,
    A.-B. Mahler, and S.~C. McClain.  \newblock First results from a
    dual photoelastic-modulator-based polarimetric camera.  \newblock
    {\em Appl. Opt.}, 49:2929--2946, 2010.

    % \bibitem{dinerDavis07} D.~J. Diner, A.~Davis, B.~Hancock,
    %   G.~Gutt, R.~A. Chipman, and B.~Cairns.  \newblock
    %   Dual-photoelastic-modulator-based polarimetric imaging concept
    %   for aerosol remote sensing.  \newblock {\em Appl. Opt.},
    %   46(35):8428--8445, 2007.

  \bibitem{diner} D.~J. Diner and J.~V. Martonchik.  \newblock
    Atmospheric transmittance from spacecraft using multiple view
    angle imagery.  \newblock {\em Appl. Opt.}, 24:3503--3511, 1985.

  \bibitem{fattal} R.~Fattal.  \newblock Single image dehazing.
    \newblock In {\em ACM SIGGRAPH}, 72:1--72:9, 2008.

  \bibitem{fuchs} C.~Fuchs, M.~Heinz, M.~Levoy, H.-P. Seidel, and
    H.~P.~A. Lensch.  \newblock Combining confocal imaging and
    descattering.  \newblock {\em Computer Graphics Forum},
    27:1245--1253, 2008.

  \bibitem{gorbunov} M.~Gorbunov, S.~Sokolovsky,
    L.~Bengtsson, %and M.-P.-I. f{\"u}r Meteorologie.
    \newblock {\em Space Refractive Tomography of the Atmosphere:
      Modeling of Direct and Inverse Problems}.  \newblock
    Max-Planck-Institut f{\"u}r Meteorologie report, 1996.

  \bibitem{gregson} J.~Gregson, M.~Krimerman, M.~B. Hullin, and
    W.~Heidrich.  \newblock Stochastic tomography and its applications
    in 3D imaging of mixing fluids.  \newblock {\em ACM TOG},
    31:52:1--52:10, 2012.

  \bibitem{hansen} J.~Hansen and L.~Travis.  \newblock Light
    scattering in planetary atmospheres.  \newblock {\em Space Science
      Reviews}, 16:527--610, 1974.

  \bibitem{he} K.~He, J.~Sun, and X.~Tang.  \newblock Single image
    haze removal using dark channel prior.  \newblock {\em IEEE
      Trans. PAMI}, 33(12):2341 --2353, 2011.

  \bibitem{horstmeyer} R.~Horstmeyer, G.~Euliss, R.~Athale, and
    M.~Levoy.  \newblock Flexible multimodal camera using a light
    field architecture.  \newblock Proc. IEEE ICCP, 2009.

  \bibitem{ihrke} I.~Ihrke, K.~N. Kutulakos, H.~P.~A. Lensch,
    M.~Magnor, and W.~Heidrich.  \newblock State of the art in
    transparent and specular object reconstruction, 2008.

  \bibitem{johnsen} S.~Johnsen and H.~M. Sosik.  \newblock Shedding
    light on light in the ocean.  \newblock {\em Oceanus}, 43:24--28,
    2004.

  \bibitem{Joshi} N. Joshi, W. Matusik, and S. Avidan.  \newblock
    Natural Video Matting using Camera Arrays.  \newblock {\em ACM
      TOG} 25, 2006

  \bibitem{Joshi2} N. Joshi, and M. Cohen.  \newblock Seeing
    Mt. Rainier: Lucky Imaging for Multi-Image Denoising, Sharpening,
    and Haze Removal.  \newblock IEEE ICCP 2010.

  \bibitem{kalashnikova} O.~V. Kalashnikova, M.~J. Garay, A.~B. Davis,
    D.~J. Diner, and J.~V.  Martonchik.  \newblock Sensitivity of
    multi-angle photo-polarimetry to vertical layering and mixing of
    absorbing aerosols: Quantifying measurement uncertainties.
    \newblock {\em J. Quantitative Spectroscopy and Radiative
      Transfer}, 112:2149 -- 2163, 2011.

  \bibitem{kim} J.~Kim, D.~Lanman, Y.~Mukaigawa, and R.~Raskar.
    \newblock Descattering transmission via angular filtering.
    \newblock In. Proc. ECCV, 86--99, 2010.

    % \bibitem{kokhan} A.~Kokhanovsky.  \newblock {\em Light
    %   Scattering Media Optics}.  \newblock Springer Praxis Books /
    %   Environmental Sciences, 2005.

  \bibitem{kopal} S.~J. Koppal, I. Gkioulekas, T. Zickler, and
    G. Barrows \newblock Wide-angle micro sensors for vision on a
    tight budget \newblock In {\em Proc. IEEE CVPR}, 2011

  \bibitem{kopf} J.~Kopf, B.~Neubert, B.~Chen, M.~F. Cohen,
    D.~Cohen-Or, O.~Deussen, M.~Uyttendaele, and D.~Lischinski.
    \newblock Deep photo: Model-based photograph enhancement and
    viewing.  \newblock {\em ACM TOG}, 27:116:1--116:10, 2008.

  \bibitem{kratz} L.~Kratz and K.~Nishino.  \newblock Factorizing
    scene albedo and depth from a single foggy image.  \newblock In
    {\em Proc. IEEE ICCV}, 2009.

  \bibitem{Levi1980} L.~Levi.  \newblock {\em Applied Optics},
    volume~2.  \newblock Jhon Wiley \& Sons, Inc, 1980.

  \bibitem{levoy} M.~Levoy, B.~Chen, V.~Vaish, M.~Horowitz,
    I.~McDowall, and M.~Bolas.  \newblock Synthetic aperture confocal
    imaging.  \newblock {\em ACM TOG}, 23(3):825--834, Aug. 2004.

    % \bibitem{levy} M.~Levy, Y.~Lehahn, J.-M. Andre, L.~Memery,
    %   H.~Loisel, and E.~Heifetz.  \newblock {Production regimes in
    %   the northeast Atlantic: A study based on Sea-viewing Wide
    %   Field-of-view Sensor (SeaWiFS) chlorophyll and ocean general
    %   circulation model mixed layer depth}.  \newblock {\em
    %   J. Geophys. Res.}, 110(C7):C07S10, 2005.

  \bibitem{Seidel} A. Lintu, L. Hoffmann, M. Magnor, H.~P.~A. Lensch,
    H. Seidel \newblock 3D Reconstruction of Reflection Nebulae from a
    Single Image \newblock Proc. Vision, Modeling, and Visualization,
    2007

  \bibitem{Martonchik2009} J.~V. Martonchik, R.~A. Kahn, and
    D.~J. Diner.  \newblock {Retrieval of aerosol properties over land
      using MISR observations}.  \newblock In A.~A. Kokhanovsky and
    G.~Leeuw, editors, {\em Satellite Aerosol Remote Sensing over
      Land}, pages 267--293, Springer Praxis Books, Berlin 2009.

  \bibitem{messer} H.~Messer, A.~Zinevich, and P.~Alpert.  \newblock
    Environmental sensor networks using existing wireless
    communication systems for rainfall and wind velocity measurements.
    \newblock {\em IEEE Instrumentation Measurement Mag.}, 15:32 --38,
    2012.

  \bibitem{Saito} D. Miyazaki, M. Saito, Y. Sato, and K. Ikeuchi.
    \newblock Determining surface orientations of transparent objects
    based on polarization degress in visible and infrared wavelengths.
    \newblock {\em JOSA A}, 19:687-694, 2002.

  \bibitem{vanMol} B.~van Mol and K.Ruddick.  \newblock The compact
    high resolution imaging spectrometer (CHRIS): the future of
    hyperspectral satellite sensors.  \newblock Airborne Imaging
    Spectros. Worksh., 2004.

  \bibitem{Marschner} J.~T. Moon, B. Walter, and S.~R. Marschner.
    \newblock Rendering Discrete Random Media Using Precomputed
    Scattering Solutions \newblock Eurographics Symposium on Rendering
    2007

    % \bibitem{moses} W.~J. Moses, A.~A. Gitelson, S.~Berdnikov, and
    %   V.~Povazhnyy.  \newblock Estimation of chlorophyll- a
    %   concentration in case ii waters using modis and meris
    %   data-successes and challenges.  \newblock {\em Environmental
    %   Research Letters}, 4:045005, 2009.

  \bibitem{narasimhan2} S.~G. Narasimhan and S.~K. Nayar.  \newblock
    Vision and the atmosphere.  \newblock {\em
      Int. J. Comput. Vision}, 48:233--254, 2002.

  \bibitem{oakley} J.~P. Oakley and B.~L. Satherley.  \newblock
    Improving image quality in poor visibility conditions using a
    physical model for contrast degradation.  \newblock {\em
      Trans. Img. Proc.}, 7:167--179, 1998.

  \bibitem{Peleg} S. Peleg, M. Ben-Ezra, and Y. Pritch.  \newblock
    OmniStereo: Panoramic Stereo Imaging \newblock {\em IEEE
      Trans. PAMI}, 279-290, 2001.

  \bibitem{Perwass} C. Perwass and L. Wietzke.  \newblock Single lens
    3D-camera with extended depth-of-field.  \newblock SPIE 8291,
    Human Vision and Electronic Imaging XVII, 2012

  \bibitem{Hschechner2} Y.~Y. Schechner, S.~G. Narasimhan, and
    S.~K. Nayar.  \newblock Polarization-based vision through haze.
    \newblock {\em Appl. Opt.}, 42:511--525, 2003.

  \bibitem{seiz} G. Seiz, E. P. Balsavias and A. Gruen.  Cloud mapping
    from the ground: use of photogrammetric methods.  {\em
      Photogrammetric Eng. Rem. Sense} 68:941-951, 2002.

  \bibitem{Snavely} \newblock N. Snavely, I. Simon, M. Goesele,
    R. Szeliski, S.~M. Seitz.  \newblock Scene Reconstruction and
    Visualization From Community Photo Collections.  {\em Proceedings
      of IEEE}, 2010.

  \bibitem{tan} R.~Tan.  \newblock Visibility in bad weather from a
    single image.  \newblock In {\em Proc. IEEE CVPR}, 2008.

  \bibitem{sun_composition} Wikipedia.  \newblock Sunlight ---
    wikipedia{,} the free encyclopedia, 2012.  \newblock [Online;
    accessed 6-August-2012].

  \bibitem{wright} T.~E. Wright, M.~Burton, D.~M. Pyle, and
    T.~Caltabiano.  \newblock {Scanning tomography of SO2 distribution
      in a volcanic gas plume}.  \newblock {\em Geophys. Res. Lett.},
    35:L17811, 2008.

  \bibitem{BFGS} C.~Zhu, R.~H. Byrd, P.~Lu, and J.~Nocedal.  \newblock
    Algorithm 778: L-BFGS-B: Fortran subroutines for large-scale
    bound-constrained optimization.  \newblock {\em ACM
      Trans. Math. Softw.}, 23:550--560, 1997.

  \bibitem{Milanfar} X. Zhu and P. Milanfar.  \newblock Removing
    atmospheric turbulence via space-invariant deconvolution.
    \newblock {\em IEEE Trans. PAMI} 35:157-170, 2013

  \end{thebibliography}

}

\end{document}

%
% David brady
% Sing Bing Kang
%
% keep article 20 and remove all other citations
%